---
title: "200514_plotHeatShock"
author: "Erin Nishimura"
date: "05/14/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Load Packages

dplyr  

ggplot2

```{r setup,  echo = FALSE, warning = FALSE, message = FALSE}

print("script initiated on", quote = FALSE)
Sys.time()

#loading packages
library(dplyr)
library(ggplot2)
library(RColorBrewer)

```


## Import

Import the data. Organize the order of factor data. Clean up the dataframe.

```{r, echo = TRUE}

# import the data
clusterTable <- read.table(file = "../1_input/figure7_sourcedata_200514.txt", header = TRUE) 


RNAsPerClusterTable <- read.table(file = "../1_input/figure7_sourcedata2_200520.txt", header = TRUE) 

```

## Exploratory Data Analysis

Explore the data. Determine the dimensions & the data structure

```{r pressure, echo=TRUE}
dim(clusterTable)
head(clusterTable)
str(clusterTable)

# Set factor levels
clusterTable$Condition <- factor(clusterTable$Condition, levels = c("No_HS", "HS"))

clusterTable$Probe <- factor(clusterTable$Probe, levels = c("gpd-2", "set-3", "B04957"))
str(clusterTable)
```


## Statistics - Percentage of clusters co-localizing with P granule markers

Calculate a set of p values determining differences between heat shock and non-heat shock conditions, this time describing differences between the metric that shows the fraction of probset-specific mRNA clusters to overlap with P granule markers.

```{r, echo=TRUE}

# Calculate p-values for each probe set
# Metric: the fraction of mRNA clusters overlapping with P granules
# Conditions: heat shock versus no heat shock conditions

# Calculate the t-test - Fract_clusters_in_Pgran
calculate_fract_pvalue <- function(data, Probeset){
  # Filter for given probeset
  transcript_quantTable <- data %>%
    filter(Probe == Probeset)
  
  # Save the p-value
  ttestresult <- t.test(Fract_clusters_in_Pgran*100 ~ Condition, data = transcript_quantTable)
  
  # Return the p-value
  return(ttestresult)
}

```

```{r, echo=FALSE}
# Calculate p values:



print("t.test calculation - gpd-2 Probe set; Fraction of gpd-2 clusters co-localizing to P granules in non-heat shock versus heat shock conditions:")
calculate_fract_pvalue(clusterTable, "gpd-2")

print("t.test calculation - set-3 Probe set; Fraction of set-3 clusters co-localizing to P granules in non-heat shock versus heat shock conditions:")
calculate_fract_pvalue(clusterTable, "set-3")

print("t.test calculation - B04957 Probe set; Fraction of B049457 clusters co-localizing to P granules in non-heat shock versus heat shock conditions:")
calculate_fract_pvalue(clusterTable, "B04957")


# Save p values:
pvalue_1 <- calculate_fract_pvalue(clusterTable, "gpd-2")$p.value
pvalue_2 <- calculate_fract_pvalue(clusterTable, "set-3")$p.value
pvalue_3 <- calculate_fract_pvalue(clusterTable, "B04957")$p.value
```

## Plot - Figure 7B - Percentage of clusters co-localizing with P granule markers

For each probeset, plot the percentage of RNA clusters that co-localize to P granules under non-heat schock and heat shock conditions.

```{r, echo=FALSE}



# gpd-2, filter & plot
plot_data <- clusterTable %>%
    filter(Probe == "gpd-2")

p1 <- ggplot(plot_data, aes(Condition, Fract_clusters_in_Pgran*100, fill=Condition)) +
  geom_boxplot(position=position_dodge(0.8), outlier.shape = NA) +
  geom_point(position=position_jitterdodge(dodge.width=0.8), alpha = 0.5) +
  ylim(c(0, 100))+
  annotate(geom="text", x=1.5, y=80, label=pvalue_1, color="#66CC66")+
  annotate(geom="text", x=1.5, y=90, label = "gpd-2")+
  scale_fill_manual(values=c("#CCCCCC", "#66CC66")) +
  theme_classic() + 
  theme(strip.text = element_text(face = "italic"),
        panel.grid.major = element_line(colour = "grey"), 
        axis.text.y = element_text(angle=90), 
        axis.text.x = element_text(angle=90))  

# set-3, filter and plot:

plot_data <- clusterTable %>%
    filter(Probe == "set-3")

p2 <- ggplot(plot_data, aes(Condition, Fract_clusters_in_Pgran*100, fill=Condition)) +
  geom_boxplot(position=position_dodge(0.8), outlier.shape = NA) +
  geom_point(position=position_jitterdodge(dodge.width=0.8), alpha = 0.5) +
  ylim(c(0, 100))+
  annotate(geom="text", x=1.5, y=80, label=pvalue_2, color="#66CC66")+
  annotate(geom="text", x=1.5, y=90, label = "set-3")+
  scale_fill_manual(values=c("#CCCCCC", "#66CC66")) +
  theme_classic() + 
  theme(strip.text = element_text(face = "italic"),
        panel.grid.major = element_line(colour = "grey"), 
        axis.text.y = element_text(angle=90), 
        axis.text.x = element_text(angle=90))  

# B04957, filter & plot: 

plot_data <- clusterTable %>%
    filter(Probe == "B04957")

p3 <- ggplot(plot_data, aes(Condition, Fract_clusters_in_Pgran*100, fill=Condition)) +
  geom_boxplot(position=position_dodge(0.8), outlier.shape = NA) +
  geom_point(position=position_jitterdodge(dodge.width=0.8), alpha = 0.5) +
  ylim(c(0, 100))+
  annotate(geom="text", x=1.5, y=80, label=pvalue_3, color="#66CC66")+
  annotate(geom="text", x=1.5, y=90, label = "B04957")+
  scale_fill_manual(values=c("#CCCCCC", "#66CC66")) +
  theme_classic() + 
  theme(strip.text = element_text(face = "italic"),
        panel.grid.major = element_line(colour = "grey"), 
        axis.text.y = element_text(angle=90), 
        axis.text.x = element_text(angle=90))  

# Display output plots

p1
p2
p3

# Save output plots to a .pdf file
date <- format(Sys.time(), "%y%m%d")
#set filename
filename <- paste("../3_output/", date, "_output_plots_Fig7B.pdf", sep = "")

#export pdf
pdf(file = filename, width = 6, height = 6)
p1
p2
p3
dev.off()
dev.off()

```

## Statistics - Raw # of clusters co-localizing with P granule markers

Now, let's switch to looking at the raw numbers of RNA clusters for each probeset.

Are there differences between no heat shock & heat shock between the number of raw RNA clusters that co-localize with P granule markers, for each probeset?

```{r, echo=TRUE}

# Calculate p-values for each probe set
# Metric: total number of mRNA clusters overlapping with P granules
# Conditions: heat shock versus no heat shock conditions

# Calculate the t-test - Fract_clusters_in_Pgran
calculate_raw_pvalue <- function(data, Probeset){
  # Filter for given probeset
  transcript_quantTable <- data %>%
    filter(Probe == Probeset)
  
  # Save the p-value
  ttestresult <- t.test(Overlap ~ Condition, data = transcript_quantTable)
  
  # Return the p-value
  return(ttestresult)
}

```


```{r, echo=FALSE}
# Calculate p values:


print("t.test calculation - gpd-2 Probe set; Raw # of gpd-2 clusters co-localizing to P granules in non-heat shock versus heat shock conditions:")
calculate_raw_pvalue(clusterTable, "gpd-2")

print("t.test calculation - set-3 Probe set; Raw # of set-3 clusters co-localizing to P granules in non-heat shock versus heat shock conditions:")
calculate_raw_pvalue(clusterTable, "set-3")

print("t.test calculation - B04957 Probe set; Raw # of B049457 clusters co-localizing to P granules in non-heat shock versus heat shock conditions:")
calculate_raw_pvalue(clusterTable, "B04957")

# Save p values:
pvalue_4 <- calculate_raw_pvalue(clusterTable, "gpd-2")$p.value
pvalue_5 <- calculate_raw_pvalue(clusterTable, "set-3")$p.value
pvalue_6 <- calculate_raw_pvalue(clusterTable, "B04957")$p.value

```


## Plot - Supplemental Figure - Raw # of clusters co-localizing with P granule markers  

For each probeset, plot the number of RNA clusters that co-localize to P granules under non-heat schock and heat shock conditions.


```{r, echo=FALSE}

# Plot boxplot
#clusterTable$Probe <- factor(clusterTable$Probe, levels = c("gpd-2", "set-3", "B04957"))

p4 <- ggplot(clusterTable, aes(Probe, Overlap, fill=Condition)) +
  geom_boxplot(position=position_dodge(0.8), outlier.shape = NA) +
  geom_point(position=position_jitterdodge(dodge.width=0.8), alpha = 0.5) +
  ylim(c(0, 60))+
  annotate(geom="text", x=1, y=58, label=pvalue_4)+
  annotate(geom="text", x=2, y=53, label=pvalue_5)+
  annotate(geom="text", x=3, y=48, label=pvalue_6)+
  scale_fill_manual(values=c("#CCCCCC", "#66CC66")) +
  theme_classic() + 
  theme(strip.text = element_text(face = "italic"),
        panel.grid.major = element_line(colour = "grey"), 
        axis.text.y = element_text(angle=90), 
        axis.text.x = element_text(angle=90))  

p4

# Export .pdf plot

#set filename
filename <- paste("../3_output/", date, "_output_plots_FigS7.pdf", sep = "")

#export pdf
pdf(file = filename, width = 8, height = 6)
p4
dev.off()
dev.off()


``` 
    
    
## Supplemental Figure - 13A - RNA cluster characteristics with and without heat shock

Create a figure that describes the phenotype of heat shock by illustrating how many clusters there are before and and after heat shock and also how many RNAs are expected to be within each cluster before and after heat shock.

Exploratory data analysis on the metrics:


```{r, echo = FALSE}


# Exploratory Data Analysis
dim(RNAsPerClusterTable)
head(RNAsPerClusterTable)


# Set factor levels
RNAsPerClusterTable$Condition <- factor(RNAsPerClusterTable$Condition, levels = c("No_HS", "HS"))

RNAsPerClusterTable$Probe <- factor(RNAsPerClusterTable$Probe, levels = c("gpd-2", "set-3", "B04957"))

str(RNAsPerClusterTable)

```

Merge and process the information
```{r, echo = FALSE}

# Calcualte the mean quantity of RNAs per cluster for each embryo
per_embryo_RNA_cluster_table <- RNAsPerClusterTable %>%
  group_by(Image_n, Probe) %>%
  mutate(number_RNAs = mean(n_rna), quantity = n()) %>%
  distinct(Image_n, .keep_all=TRUE)  

dim(per_embryo_RNA_cluster_table)

# Calculate the mean RNAs per cluster for each condition & probeset
# Calcualte the mean number of clusters for each condition & probeset

per_Condition_RNA_cluster_table <- per_embryo_RNA_cluster_table %>%
  group_by(Condition, Probe) %>%
  mutate(mean_RNAs = mean(number_RNAs), se_RNA = sd(number_RNAs)/sqrt(number_RNAs), number_clusters = mean(quantity), sd_clusters = sd(quantity)) %>%
  distinct(Condition, .keep_all=TRUE)  


```


## Plot Figure S13A

```{r, echo = FALSE}


plot_S13A <- ggplot(per_Condition_RNA_cluster_table, aes(number_clusters, mean_RNAs, color=Probe))+
  geom_point( aes(shape=Condition, color=Probe), size = 6) + 
  geom_segment(aes(x = number_clusters, y = (mean_RNAs-se_RNA), xend= number_clusters, yend= (mean_RNAs+se_RNA)))+  
  geom_segment(aes(x = (number_clusters-sd_clusters), y = mean_RNAs, xend= (number_clusters+sd_clusters), yend= mean_RNAs))+
  ylim(c(5, 10))+
  xlab("Number of clusters") + 
  ylab("Number of RNAs per cluster")+
  scale_color_manual(values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  scale_shape_manual(values=c(19, 17))+
  theme_classic() + 
  theme(strip.text = element_text(face = "italic"),
        panel.grid.major = element_line(colour = "grey"), 
        axis.text.y = element_text(angle=90), 
        axis.text.x = element_text(angle=90))    
  
plot_S13A


```

```{r, echo = FALSE}

# Save plot

# Export .pdf plot

#set filename
filename <- paste("../3_output/", date, "_output_plots_FigS13A.pdf", sep = "")

#export pdf
pdf(file = filename, width = 8, height = 6, useDingbats=FALSE)
plot_S13A
dev.off()
dev.off()

```

## Session info

```{r, echo = TRUE}
sessionInfo()
```
