---
title: "191029_RBP_contributions_granules"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Load Packages

dplyr  

ggplot2

ggthemes

gridExtra

tidyr

```{r setup,  echo = FALSE, warning = FALSE, message = FALSE}

print("script initiated on", quote = FALSE)
Sys.time()

#loading packages
library(dplyr)
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(tidyr)

```

## Import

Import the data. Organize the order of factor data. Clean up the dataframe.

```{r, echo = FALSE}

# import the data
quantTable <- read.table(file = "../1_input/figure5_sourcedata1_191101.txt", header = TRUE) 

# re-order the factors
quantTable$Cell_Stage <- factor(quantTable$Cell_Stage, levels = c("4-cell", "8-cell", "12-14-cell", "16-24-cell", "26-48-cell"))

```

## Exploratory Data Analysis

Explore the data. Determine the dimensions & the data structure

```{r pressure, echo=TRUE}
dim(quantTable)
head(quantTable)
str(quantTable)

```

## Data filtering

I don't really want to include the pie-1 data at this point. 

```{r, echo = FALSE, warning = FALSE}

#Filter out pie-1 values
noPie <- filter(quantTable, RNAi_Condition != "pie-1")

#Reset the order of factors for RNAi conditions
noPie$RNAi_Condition <- factor(noPie$RNAi_Condition, levels = c("L4440", "oma-1", "mex-3", "spn-4", "spn-4_mex-3"))
```



## Functionifying 

I wrote two functions to parse through the data, one will be for reporting means of count data. The other will be for log transforming median data.


```{r, echo = FALSE, warning = FALSE, message = FALSE}
# Make a function to summarize the data
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

# median relative
data_summarylog <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(median = log(median(x[[col]], na.rm=TRUE),10),
      sd = sd(log(x[[col]],10), na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("median" = varname))
  data_sum[is.na(data_sum)] <- 0
 return(data_sum)
}


# Make a plotting function
# Make a function to plot the data
# plotStats(dataset, varname, titleY)
#    dataset <- this is a dataset in the form produced by 'data_summary'
#    varname <- this corresponds to a colname in 'dataset'
#    titleY  <- this is a title for the Y axis of the plot
plotStats <- function(dataset, varname, titleY){
  ggplot(dataset, aes(x=Cell_Stage, y=varname, group=RNAi_Condition, color=RNAi_Condition, shape = RNAi_Condition))+ 
  geom_ribbon(aes(ymin=varname-sd, ymax=varname+sd, fill = RNAi_Condition), alpha=0.1, linetype = 0) +
  facet_grid(. ~ Transcript , scales = "free_y") +
  labs(y = titleY) +
  geom_line() + geom_point() +
  theme_linedraw() +
  scale_colour_colorblind() +
  scale_fill_colorblind() +
  scale_shape_manual(values=c(15, 16, 17, 18, 6)) +
  theme(strip.text = element_text(face = "italic"), 
        panel.grid.minor = element_blank(), 
        panel.grid.major = element_line(colour = "grey"), 
        axis.text.y = element_text(angle=90), 
        axis.text.x = element_text(angle=90))
}


```


## Plotting

### Plot the total numbers of RNA molecules quantified
```{r, echo = FALSE, warning = FALSE, message = FALSE}

#Summarize the data for Number of RNAs
noPiel_summ_total_RNA <- data_summary(noPie, varname="Total_No_RNA", 
                    groupnames=c("Transcript", "RNAi_Condition", "Cell_Stage"))

# plot number of RNA molecules by transcript type
p_rnaTot <- plotStats(noPiel_summ_total_RNA, noPiel_summ_total_RNA$Total_No_RNA, "Total number of mRNA molecules")

p_rnaTot

```

### Plot the log number of RNA clusters

```{r, echo = FALSE, warning = FALSE, message = FALSE}


#Summarize the data for Number of clusters
noPiel_total_clusters <- data_summary(noPie, varname="No_RNA_clusters", 
                    groupnames=c("Transcript", "RNAi_Condition", "Cell_Stage"))

# plot number of clusters by transcript type 

p_clustTot <- plotStats(noPiel_total_clusters, noPiel_total_clusters$No_RNA_clusters, "Number of RNA clusters")

p_clustTot 

```


### The fraction of mRNA in clusters (versus as singletons)

```{r, echo = FALSE, warning = FALSE, message = FALSE}

noPie <- noPie %>%
  mutate(fraction2 = No_RNA_in_clusters/Total_No_RNA)

#Summarize the data for Fraction of mRNAs in clusters
noPiel_fraction_inclusters <- data_summary(noPie, varname="fraction2", 
                    groupnames=c("Transcript", "RNAi_Condition", "Cell_Stage"))

# plot number of RNA molecules by transcript type
p_fraction <- plotStats(noPiel_fraction_inclusters, noPiel_fraction_inclusters$fraction2, "fraction mRNAs in clusters")

p_fraction

quantTable <- quantTable %>%
  mutate(fraction2 = No_RNA_in_clusters/Total_No_RNA)


```

### The average mRNA molecules per cluster

```{r, echo = FALSE, warning = FALSE, message = FALSE}

#Summarize the data for Average mRNA molecules per cluster
noPie_Average_mR_per_C <- data_summary(noPie, varname="Average_No_RNA_per_Cluster", 
                    groupnames=c("Transcript", "RNAi_Condition", "Cell_Stage"))

# plot Average mRNA molecules per cluster by transcript type

p_Average <- plotStats(noPie_Average_mR_per_C, noPie_Average_mR_per_C$Average_No_RNA_per_Cluster, "Average number of mRNAs per cluster")

p_Average 

```
    
    
## Plot figures

Export plots to files
```{r, fig.path = '../3_output', echo = FALSE, warning = FALSE}
#Set date
date <- format(Sys.time(), "%y%m%d")
#set filename
filename <- paste("../3_output/", date, "_output_plots_Fig5c.pdf", sep = "")

#export pdf
pdf(file = filename, width = 8, height = 6)
grid.arrange(p_rnaTot, p_Average, ncol = 1)
dev.off()
dev.off()

#Report
print("The following output plot was saved: ", quote = FALSE)
print(filename, quote = FALSE)

```

## Supplemental figures

Export plots to be used for a supplemental Figure

```{r, fig.path = '../3_output', echo = FALSE, warning = FALSE}
#Set date
date <- format(Sys.time(), "%y%m%d")
#set filename
filename <- paste("../3_output/", date, "_output_plots_SuppFig5.pdf", sep = "")

#export pdf
pdf(file = filename, width = 8, height = 12)
grid.arrange(p_rnaTot, p_clustTot, p_fraction, p_Average, ncol = 1)
dev.off()
dev.off()

#Report
print("The following output plot was saved: ", quote = FALSE)
print(filename, quote = FALSE)

```


## Statistics

Calculate statistical differences

```{r, echo = FALSE, warning = FALSE}

t.testdiffs_molecules <- function(data, transcript, RBP){
  # select transcript & RBP
  transcript_quantTable <- data%>%
    filter(Transcript == transcript) %>%
    filter(RNAi_Condition == "L4440" | RNAi_Condition == RBP)
  
  #Adjust factor levels
  transcript_quantTable$RNAi_Condition <- factor(transcript_quantTable$RNAi_Condition, levels = c("L4440", RBP))
  
  # summarize transcript-specific data
  transcript_SummaryTable <- transcript_quantTable%>%
    dplyr::group_by(Cell_Stage, RNAi_Condition) %>%
    dplyr::summarize(value = list(Total_No_RNA)) %>%
    spread(RNAi_Condition, value)
  
  # Perform t-tests
  transcript_pvalues <- data.frame()

  for (i in 1:5){
  transcript_pvalues[i,1] <- t.test(transcript_SummaryTable[[i,2]], transcript_SummaryTable[[i,3]])$p.value
    rownames(transcript_pvalues)[i] <- as.character(transcript_SummaryTable$Cell_Stage[i])
  }
  colnames(transcript_pvalues) <- c("p.value")
  
  # Perform multiple test corrections using the Benjamini Hochberg method
  transcript_pvalues <- cbind(transcript_pvalues, BH_padjust = p.adjust(transcript_pvalues$p.value, method = "BH"))

  # Report p values
  return(transcript_pvalues)
}

t.testdiffs_clusters <- function(data, transcript, RBP){
  # select transcript & RBP
  transcript_quantTable <- data%>%
    filter(Transcript == transcript) %>%
    filter(RNAi_Condition == "L4440" | RNAi_Condition == RBP)
  
  #Adjust factor levels
  transcript_quantTable$RNAi_Condition <- factor(transcript_quantTable$RNAi_Condition, levels = c("L4440", RBP))
  
  # summarize transcript-specific data
  transcript_SummaryTable <- transcript_quantTable%>%
    dplyr::group_by(Cell_Stage, RNAi_Condition) %>%
    dplyr::summarize(value = list(No_RNA_clusters)) %>%
    spread(RNAi_Condition, value)
  
  # Perform t-tests
  transcript_pvalues <- data.frame()

  for (i in 1:5){
  transcript_pvalues[i,1] <- t.test(transcript_SummaryTable[[i,2]], transcript_SummaryTable[[i,3]])$p.value
    rownames(transcript_pvalues)[i] <- as.character(transcript_SummaryTable$Cell_Stage[i])
  }
  colnames(transcript_pvalues) <- c("p.value")
  
  # Perform multiple test corrections using the Benjamini Hochberg method
  transcript_pvalues <- cbind(transcript_pvalues, BH_padjust = p.adjust(transcript_pvalues$p.value, method = "BH"))

  # Report p values
  return(transcript_pvalues)
}


t.testdiffs_fraction <- function(data, transcript, RBP){
  # select transcript & RBP
  transcript_quantTable <- data%>%
    filter(Transcript == transcript) %>%
    filter(RNAi_Condition == "L4440" | RNAi_Condition == RBP)
  
  #Adjust factor levels
  transcript_quantTable$RNAi_Condition <- factor(transcript_quantTable$RNAi_Condition, levels = c("L4440", RBP))
  
  # summarize transcript-specific data
  transcript_SummaryTable <- transcript_quantTable%>%
    dplyr::group_by(Cell_Stage, RNAi_Condition) %>%
    dplyr::summarize(value = list(fraction_RNA_in_clusters)) %>%
    spread(RNAi_Condition, value)
  
  # Perform t-tests
  transcript_pvalues <- data.frame()

  for (i in 1:5){
  transcript_pvalues[i,1] <- t.test(transcript_SummaryTable[[i,2]], transcript_SummaryTable[[i,3]])$p.value
    rownames(transcript_pvalues)[i] <- as.character(transcript_SummaryTable$Cell_Stage[i])
  }
  colnames(transcript_pvalues) <- c("p.value")
  
  # Perform multiple test corrections using the Benjamini Hochberg method
  transcript_pvalues <- cbind(transcript_pvalues, BH_padjust = p.adjust(transcript_pvalues$p.value, method = "BH"))

  # Report p values
  return(transcript_pvalues)
}


t.testdiffs_avg <- function(data, transcript, RBP){
  # select transcript & RBP
  transcript_quantTable <- data%>%
    filter(Transcript == transcript) %>%
    filter(RNAi_Condition == "L4440" | RNAi_Condition == RBP)
  
  #Adjust factor levels
  transcript_quantTable$RNAi_Condition <- factor(transcript_quantTable$RNAi_Condition, levels = c("L4440", RBP))
  
  # summarize transcript-specific data
  transcript_SummaryTable <- transcript_quantTable%>%
    dplyr::group_by(Cell_Stage, RNAi_Condition) %>%
    dplyr::summarize(value = list(Average_No_RNA_per_Cluster)) %>%
    spread(RNAi_Condition, value)
  
  # Perform t-tests
  transcript_pvalues <- data.frame()

  for (i in 1:5){
  transcript_pvalues[i,1] <- t.test(transcript_SummaryTable[[i,2]], transcript_SummaryTable[[i,3]])$p.value
    rownames(transcript_pvalues)[i] <- as.character(transcript_SummaryTable$Cell_Stage[i])
  }
  colnames(transcript_pvalues) <- c("p.value")
  
  # Perform multiple test corrections using the Benjamini Hochberg method
  transcript_pvalues <- cbind(transcript_pvalues, BH_padjust = p.adjust(transcript_pvalues$p.value, method = "BH"))

  # Report p values
  return(transcript_pvalues)
}


```

### Statistical significance testing for the number of total mRNA molecules


```{r, echo = FALSE}
#### MOLECULES #####

## Test molecules p-values for chs-1
chs1_molecules_pvals <- lapply(c("oma-1", "mex-3", "spn-4", "spn-4_mex-3"), 
                               function(x) {t.testdiffs_molecules(quantTable, "chs-1", x)})
names(chs1_molecules_pvals) <- c("oma-1", "mex-3", "spn-4", "spn-4_mex-3")
print("number of molecules = chs-1")
print(chs1_molecules_pvals)

# p <- chs1_molecules_pvals[[4]][[2]]
# (0.05 < p )
# (0.005 < p )
# (0.0005 < p )
# (0.00005 < p )
# (0.000005 < p )

## Test molecules p-values for clu-1
clu_molecules_pvals <- lapply(c("oma-1", "mex-3", "spn-4", "spn-4_mex-3"), 
                               function(x) {t.testdiffs_molecules(quantTable, "clu-1", x)})
names(clu_molecules_pvals) <- c("oma-1", "mex-3", "spn-4", "spn-4_mex-3")
print("number of molecules = clu-1")
print(clu_molecules_pvals)

# p <- clu_molecules_pvals[[4]][[2]]
# (0.05 < p )
# (0.005 < p )
# (0.0005 < p )
# (0.00005 < p )
# (0.000005 < p )

## Test molecules p-values for cpg-2
cpg_molecules_pvals <- lapply(c("oma-1", "mex-3", "spn-4", "spn-4_mex-3"), 
                               function(x) {t.testdiffs_molecules(quantTable, "cpg-2", x)})
names(cpg_molecules_pvals) <- c("oma-1", "mex-3", "spn-4", "spn-4_mex-3")
print("number of molecules = cpg-2")
print(cpg_molecules_pvals)

# p <- cpg_molecules_pvals[[4]][[2]]
# (0.05 < p )
# (0.005 < p )
# (0.0005 < p )
# (0.00005 < p )
# (0.000005 < p )


## Test molecules & nos-2 p-values
nos2_molecules_pvals <- lapply(c("oma-1", "mex-3", "spn-4", "spn-4_mex-3"), 
                               function(x) {t.testdiffs_molecules(quantTable, "nos-2", x)})
names(nos2_molecules_pvals) <- c("oma-1", "mex-3", "spn-4", "spn-4_mex-3")
print("number of molecules = nos-2")
print(nos2_molecules_pvals)

# p <- nos2_molecules_pvals[[4]][[2]]
# (0.05 < p )
# (0.005 < p )
# (0.0005 < p )
# (0.00005 < p )
# (0.000005 < p )


```

### Statistical significance testing for the average number of molecules per cluster for each embryo

```{r, echo = FALSE}
#### AVERAGES #####

## Test pvalues on averages & chs-1 p-values
chs_averages_pvals <- lapply(c("oma-1", "mex-3", "spn-4", "spn-4_mex-3"), 
                               function(x) {t.testdiffs_avg(quantTable, "chs-1", x)})
names(chs_averages_pvals) <- c("oma-1", "mex-3", "spn-4", "spn-4_mex-3")
print("average molecules per cluster = chs-1")
print(chs_averages_pvals)

p <- chs_averages_pvals[[4]][[2]]

# (0.05 > p )
# (0.005 > p )
# (0.0005 > p )
# (0.00005 > p )
# (0.000005 > p )
# # (0.05 < 2.751155e-02)
# # (0.005 < 2.751155e-02)
# # (0.0005 < 2.751155e-02)
# # (0.00005 < 2.751155e-02)


## Test averages & clu-1 p-values
clu_averages_pvals <- lapply(c("oma-1", "mex-3", "spn-4", "spn-4_mex-3"), 
                               function(x) {t.testdiffs_avg(quantTable, "clu-1", x)})
names(clu_averages_pvals) <- c("oma-1", "mex-3", "spn-4", "spn-4_mex-3")
print("average molecules per cluster = clu-1")
print(clu_averages_pvals)

# p <- clu_averages_pvals[[1]][[2]]
# 
# (0.05 < p )
# (0.005 < p )
# (0.0005 < p )
# (0.00005 < p )
# (0.000005 < p )


## Test averages & clu-1 p-values
cpg_averages_pvals <- lapply(c("oma-1", "mex-3", "spn-4", "spn-4_mex-3"), 
                               function(x) {t.testdiffs_avg(quantTable, "cpg-2", x)})
names(cpg_averages_pvals) <- c("oma-1", "mex-3", "spn-4", "spn-4_mex-3")
print("average molecules per cluster = cpg-2")
print(cpg_averages_pvals)

# p <- cpg_averages_pvals[[4]][[2]]
# 
# (0.05 < p )
# (0.005 < p )
# (0.0005 < p )
# (0.00005 < p )
# (0.000005 < p )

## Test averages & nos-2 p-values
nos2_averages_pvals <- lapply(c("oma-1", "mex-3", "spn-4", "spn-4_mex-3"), 
                               function(x) {t.testdiffs_avg(quantTable, "nos-2", x)})
names(nos2_averages_pvals) <- c("oma-1", "mex-3", "spn-4", "spn-4_mex-3")
print("average molecules per cluster = nos-2")
print(nos2_averages_pvals)

# p <- nos2_averages_pvals[[4]][[2]]
# (0.05 < p )
# (0.005 < p )
# (0.0005 < p )
# (0.00005 < p )
# (0.000005 < p )
```

```{r, echo = FALSE}
# # select nos-2 data
# nos2QuantTable <- quantTable%>%
#   filter(Transcript == "nos-2") %>%
#   filter(RNAi_Condition == "L4440" | RNAi_Condition == "spn-4_mex-3")
# 
# nos2QuantTable$RNAi_Condition <- factor(nos2QuantTable$RNAi_Condition, levels = c("L4440", "spn-4_mex-3"))
# 
# # summarize chs-1 data
# nos2SummaryTable <- nos2QuantTable%>%
#   dplyr::group_by(Cell_Stage, RNAi_Condition) %>%
#   dplyr::summarize(value = list(Average_No_RNA_per_Cluster)) %>%
#   spread(RNAi_Condition, value)
# 
# nos2SummaryTable
# # Perform t-tests
# nos2_pvalues <- data.frame()
# 
# for (i in 1:5){
#   nos2_pvalues[i,1] <- t.test(nos2SummaryTable[[i,2]], nos2SummaryTable[[i,3]])$p.value
#   rownames(nos2_pvalues)[i] <- as.character(nos2SummaryTable$Cell_Stage[i])
# }
# 
# colnames(nos2_pvalues) <- c("p.value")
# nos2_pvalues
# # Perform multiple test corrections using the Benjamini Hochberg method
# nos2_pvalues <- cbind(nos2_pvalues, BH_padjust = p.adjust(nos2_pvalues$p.value, method = "BH"))
# BH_padjust = p.adjust(nos2_pvalues$p.value, method = "BH")
# 
# # Report p values
# nos2_pvalues

```

```{r, echo = FALSE}

# quantTable$RNAi_Condition
# nos2QuantTable <- t.testdiffs(quantTable, "chs-1", "spn-4", 10)
# dim(quantTable)
# 
# # select nos-2 data
# nos2QuantTable <- quantTable%>%
#   filter(Transcript == "chs-1") %>%
#   filter(RNAi_Condition == "L4440" | RNAi_Condition == "mex-3" | RNAi_Condition == "spn-4")
# 
# nos2QuantTable$RNAi_Condition <- factor(nos2QuantTable$RNAi_Condition, levels = c("L4440", "mex-3", "spn-4"))
# 
# # summarize chs-1 data
# nos2SummaryTable <- nos2QuantTable%>%
#   dplyr::group_by(Cell_Stage, RNAi_Condition) %>%
#   dplyr::summarize(value = list(Total_No_RNA)) %>%
#   spread(RNAi_Condition, value)
# nos2SummaryTable
# mean(unlist(nos2SummaryTable[1,2]))
# mean(unlist(nos2SummaryTable[1,3]))
# mean(unlist(nos2SummaryTable[1,4]))
# 
# mean(unlist(nos2SummaryTable[3,2]))
# mean(unlist(nos2SummaryTable[3,3]))
# mean(unlist(nos2SummaryTable[3,4]))
# 
# 
# 
# unlist(nos2SummaryTable[2,3])
# unlist(nos2SummaryTable[3,3])
# # Perform t-tests
# nos2_pvalues <- data.frame()
# 
# for (i in 1:5){
#   nos2_pvalues[i,1] <- t.test(nos2SummaryTable[[i,2]], nos2SummaryTable[[i,3]])$p.value
#   rownames(nos2_pvalues)[i] <- as.character(nos2SummaryTable$Cell_Stage[i])
# }
# 
# colnames(nos2_pvalues) <- c("p.value")
# nos2_pvalues 
# # Perform multiple test corrections using the Benjamini Hochberg method
# nos2_pvalues <- cbind(nos2_pvalues, BH_padjust = p.adjust(nos2_pvalues$p.value, method = "BH"))
# BH_padjust = p.adjust(nos2_pvalues$p.value, method = "BH")
# 
# # Report p values
# nos2_pvalues


```

## Get versions
```{r, echo = TRUE}
sessionInfo()
```
