---
title: "191021_Pgran_quantification.Rmd"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Load libraries

Dependencies:   

dplyr   

ggplot2

gridExtra  

tidyr

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(gridExtra)
library(tidyr)
```

### Import

Import the data. Organize the order of factor data. Clean up the dataframe.

```{r, echo = FALSE}

# Import
quantTable <- read.table(file = "../1_input/figure2_sourcedata3_191018.txt", header = TRUE, row.names = 1) # import the data

print("Input data is ../1_input/figure2_sourceData3_191018.txt", quote = FALSE)
# re-order the factors
quantTable$Cell_Stage <- factor(quantTable$Cell_Stage, levels = c("2-cell", "4-cell", "8-cell", "12-14-cell", "16-24-cell", "26-48-cell"))


```

### Exploratory Data Analysis

Explore the data. Determine the dimensions & the data structure

```{r pressure, echo=FALSE}

# report structure
str(quantTable)
dim(quantTable)
head(quantTable)

```


# Plotting

### Ribbon plots


#### Create functions
```{r, echo = FALSE, warning = FALSE, message = FALSE}

# Make a function to summarize the data
# data_summary(data, varname, groupnames)
#    data <- this is a dataset in the form 'quantTable'
#    varname <- this corresponds to a colname in 'data'
#    groupnames  <- this is a vector of colnames in 'data' that are the factor variables.
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

# Make a function to plot the data
# plotStats(dataset, varname, titleY)
#    dataset <- this is a dataset in the form produced by 'data_summary'
#    varname <- this corresponds to a colname in 'dataset'
#    titleY  <- this is a title for the Y axis of the plot
plotStats <- function(dataset, varname, titleY){
  ggplot(dataset, aes(x=Cell_Stage, y=varname, group=Transcript, color=Transcript))+ 
  geom_ribbon(aes(ymin=varname-sd, ymax=varname+sd, fill = Transcript), alpha=0.1, linetype = 0) +
  facet_grid(. ~ Transcript , scales = "free_y") +
  labs(y = titleY) +
  geom_line() + geom_point() +
  theme_linedraw() +
  theme(strip.text = element_text(face = "italic"), 
        panel.grid.minor = element_blank(), 
        panel.grid.major = element_line(colour = "grey"), 
        axis.text.y = element_text(angle=90), 
        axis.text.x = element_text(angle=90))
}

```

#### The number of RNA molecules by transcript type
```{r, echo = FALSE, warning = FALSE, message = FALSE}
# Gather the number of total RNAs by transcript type
Pgran_summ_total_RNA <- data_summary(quantTable, varname="Total_No_RNA", 
                    groupnames=c("Transcript", "Cell_Stage"))

# plot number of RNA molecules by transcript type
p_NoRNA <- plotStats(Pgran_summ_total_RNA, Pgran_summ_total_RNA$Total_No_RNA, "Molecules of mRNA")

p_NoRNA

```


#### The number of clusters by transcript type

```{r, echo = FALSE, warning = FALSE, message = FALSE}

# plot number of clusters by transcript type
Pgran_summ_total_clusters <- data_summary(quantTable, varname="No_RNA_clusters", 
                    groupnames=c("Transcript", "Cell_Stage"))


# plot number of clusters by transcript type
p_clusters <- plotStats(Pgran_summ_total_clusters, Pgran_summ_total_clusters$No_RNA_clusters, "Clusters of mRNA")

p_clusters

```

#### The fraction of RNAs in clusters by transcript type

```{r, echo = FALSE, warning = FALSE, message = FALSE}

# calculate fraction of mRNA in clusters by transcript type
Pgran_summ_fraction <- data_summary(quantTable, varname="fraction_RNA_in_clusters",
                    groupnames=c("Transcript", "Cell_Stage"))

# plot fraction of mRNA in clusters transcript type
p_fraction <- plotStats(Pgran_summ_fraction, Pgran_summ_fraction$fraction_RNA_in_clusters, "Fraction of mRNA in clusters")

p_fraction

```


#### The RNA content of clusters by transcript type

```{r, echo = FALSE, warning = FALSE, message = FALSE}


# plot RNA content of clusters by transcript type
Pgran_avg_content <- data_summary(quantTable,
                                         varname="Average_No_RNA_per_Cluster",
                                         groupnames=c("Transcript", "Cell_Stage"))

# plot RNA content of clusters transcript type
p_content <- plotStats(Pgran_avg_content, Pgran_avg_content$Average_No_RNA_per_Cluster, "Average number of mRNAs per cluster")

p_content

```

#### Printout the pdf file output
```{r, echo = FALSE, warning = FALSE}
#Set date
date <- format(Sys.time(), "%y%m%d")
#set filename
filename <- paste("../3_output/", date, "_output_plots_Fig2c.pdf", sep = "")

#export pdf
pdf(file = filename, width = 8, height = 11)
grid.arrange(p_NoRNA, p_clusters, p_fraction, p_content, ncol = 1)
dev.off()
dev.off()

#Report
print("The following output plot was saved: ", quote = FALSE)
print(filename, quote = FALSE)

```

## Statistics

Calculate statistically significant differences

Create statistical test functions first
```{r, echo = FALSE}

t.testdiffs_molecules <- function(data, transcpt){
  
  # filter for transcripts
  transcript_quantTable <- quantTable%>%
    filter(Transcript == transcpt | Transcript == "set-3")
  
  # re-do factoring
  transcript_quantTable$Transcript <- factor(transcript_quantTable$Transcript, levels = c(transcpt, "set-3"))
  
  # summarize transcript-specific data
  transcript_SummaryTable <- transcript_quantTable%>%
    dplyr::group_by(Cell_Stage, Transcript) %>%
    dplyr::summarize(value = list(Total_No_RNA)) %>%
    spread(Transcript, value)

  # Perform t-tests
  transcript_pvalues <- data.frame()

  for (i in 1:6){
  transcript_pvalues[i,1] <- t.test(transcript_SummaryTable[[i,2]],
                                    transcript_SummaryTable[[i,3]])$p.value
    rownames(transcript_pvalues)[i] <- as.character(transcript_SummaryTable$Cell_Stage[i])
  }
  
  colnames(transcript_pvalues) <- c("p.value")
  
  # Perform multiple test corrections using the Benjamini Hochberg method
  transcript_pvalues <- cbind(transcript_pvalues, BH_padjust = p.adjust(transcript_pvalues$p.value, method = "BH"))
  
  # Report p values
  return(transcript_pvalues)
}

t.testdiffs_clusters <- function(data, transcpt){
  
  # filter for transcripts
  transcript_quantTable <- quantTable%>%
    filter(Transcript == transcpt | Transcript == "set-3")
  
  # re-do factoring
  transcript_quantTable$Transcript <- factor(transcript_quantTable$Transcript, levels = c(transcpt, "set-3"))
  # summarize transcript-specific data
  transcript_SummaryTable <- transcript_quantTable%>%
    dplyr::group_by(Cell_Stage, Transcript) %>%
    dplyr::summarize(value = list(No_RNA_clusters)) %>%
    spread(Transcript, value)

  # Perform t-tests
  transcript_pvalues <- data.frame()

  for (i in 1:6){
  transcript_pvalues[i,1] <- t.test(transcript_SummaryTable[[i,2]],
                                    transcript_SummaryTable[[i,3]])$p.value
    rownames(transcript_pvalues)[i] <- as.character(transcript_SummaryTable$Cell_Stage[i])
  }
  
  colnames(transcript_pvalues) <- c("p.value")
  
  # Perform multiple test corrections using the Benjamini Hochberg method
  transcript_pvalues <- cbind(transcript_pvalues, BH_padjust = p.adjust(transcript_pvalues$p.value, method = "BH"))
  
  # Report p values
  return(transcript_pvalues)
}

t.testdiffs_fraction <- function(data, transcpt){
  
  # filter for transcripts
  transcript_quantTable <- quantTable%>%
    filter(Transcript == transcpt | Transcript == "set-3")
  
  # re-do factoring
  transcript_quantTable$Transcript <- factor(transcript_quantTable$Transcript, levels = c(transcpt, "set-3"))
  # summarize transcript-specific data
  transcript_SummaryTable <- transcript_quantTable%>%
    dplyr::group_by(Cell_Stage, Transcript) %>%
    dplyr::summarize(value = list(fraction_RNA_in_clusters)) %>%
    spread(Transcript, value)

  # Perform t-tests
  transcript_pvalues <- data.frame()

  for (i in 1:6){
  transcript_pvalues[i,1] <- t.test(transcript_SummaryTable[[i,2]],
                                    transcript_SummaryTable[[i,3]])$p.value
    rownames(transcript_pvalues)[i] <- as.character(transcript_SummaryTable$Cell_Stage[i])
  }
  
  colnames(transcript_pvalues) <- c("p.value")
  
  # Perform multiple test corrections using the Benjamini Hochberg method
  transcript_pvalues <- cbind(transcript_pvalues, BH_padjust = p.adjust(transcript_pvalues$p.value, method = "BH"))
  
  # Report p values
  return(transcript_pvalues)
}


t.testdiffs_avg <- function(data, transcpt){
  
  # filter for transcripts
  transcript_quantTable <- quantTable%>%
    filter(Transcript == transcpt | Transcript == "set-3")
  
  # re-do factoring
  transcript_quantTable$Transcript <- factor(transcript_quantTable$Transcript, levels = c(transcpt, "set-3"))
  # summarize transcript-specific data
  transcript_SummaryTable <- transcript_quantTable%>%
    dplyr::group_by(Cell_Stage, Transcript) %>%
    dplyr::summarize(value = list(Average_No_RNA_per_Cluster)) %>%
    spread(Transcript, value)

  # Perform t-tests
  transcript_pvalues <- data.frame()

  for (i in 1:5){
  transcript_pvalues[i,1] <- t.test(transcript_SummaryTable[[i,2]],
                                    transcript_SummaryTable[[i,3]])$p.value
    rownames(transcript_pvalues)[i] <- as.character(transcript_SummaryTable$Cell_Stage[i])
  }

  colnames(transcript_pvalues) <- c("p.value")

  # Perform multiple test corrections using the Benjamini Hochberg method
  transcript_pvalues <- cbind(transcript_pvalues, BH_padjust = p.adjust(transcript_pvalues$p.value, method = "BH"))
  
  # Report p values
  return(transcript_pvalues)
}



```

### Execute the tests

```{r, echo = FALSE}

## Test molecules p-values 
molecules_pvals <- lapply(c("chs-1", "clu-1", "cpg-2", "nos-2"), 
                               function(x) {t.testdiffs_molecules(quantTable, x)})
names(molecules_pvals) <- c("chs-1", "clu-1", "cpg-2", "nos-2")
print("P-values: Number of mRNA molecules as compared to set-3" )
print(molecules_pvals)

# p <- 3.040104e-02
# (0.05 < p )
# (0.005 < p )
# (0.0005 < p )
# (0.00005 < p )
# (0.000005 < p )

## Test clusters p-values 
clusters_pvals <- lapply(c("chs-1", "clu-1", "cpg-2", "nos-2"), 
                               function(x) {t.testdiffs_clusters(quantTable, x)})
names(clusters_pvals) <- c("chs-1", "clu-1", "cpg-2", "nos-2")
print("P-values: Number of clusters as compared to set-3" )
print(clusters_pvals)

# clusters_pvals[[4]]
# p <- 0.0066173217
# (0.05 < p )
# (0.005 < p )
# (0.0005 < p )
# (0.00005 < p )
# (0.000005 < p )


## Test fraction of mRNAs in clusters  p-values 
fraction_pvals <- lapply(c("chs-1", "clu-1", "cpg-2", "nos-2"), 
                               function(x) {t.testdiffs_fraction(quantTable, x)})
names(fraction_pvals) <- c("chs-1", "clu-1", "cpg-2", "nos-2")
print("P-values: Fraction of mRNA molecules in clusters as compared to set-3" )
print(fraction_pvals)

# fraction_pvals[[4]]
# p <- 2.760786e-03
# (0.05 < p )
# (0.005 < p )
# (0.0005 < p )
# (0.00005 < p )
# (0.000005 < p )



## Test average number of mRNAs in clusters - p-values 
avgClustSize_pvals <- lapply(c("chs-1", "clu-1", "cpg-2", "nos-2"), 
                               function(x) {t.testdiffs_avg(quantTable, x)})
names(avgClustSize_pvals) <- c("chs-1", "clu-1", "cpg-2", "nos-2")
print("P-values: Average number of mRNAs in clusters as compared to set-3" )
print(avgClustSize_pvals)

# avgClustSize_pvals[[4]]
# p <- 0.0016355690
# (0.05 < p )
# (0.005 < p )
# (0.0005 < p )
# (0.00005 < p )
# (0.000005 < p )


```


# Get versions
```{r, echo = TRUE}
sessionInfo()
```
