---
title: "191022_plot_Fig4FH_gran3UTR.Rmd"
output: html_document
---

### Setting up libraries and obtaining the data

Requires:

ggplot2

ggthemes

tidyr 

broom

gridExtra

dplyr

```{r setup, include=FALSE}

print("script initiated on", quote = FALSE)
Sys.time()

library(ggplot2)
library(ggthemes)
library(gridExtra)
library(tidyr)
library(broom)
library(dplyr)

```

### Import

Import the data. Organize the order of factor data. Clean up the dataframe.

```{r, echo = FALSE}

# import the data
quantTable <- read.table(file = "../1_input/figure4_Sourcedata3_191022.txt", header = TRUE, row.names = 1) 

```

### Exploratory Data Analysis

Explore the data. Determine the dimensions & the data structure

```{r pressure, echo=TRUE}
quantTable$Cell_Stage <- factor(quantTable$Cell_Stage, levels = c("2-cell", "4-cell", "8-cell", "10-16-cell", "18-28-cell"))

dim(quantTable)
head(quantTable)
str(quantTable)
```



### Functionify the data

Write a function called data_summary to summarize the data

```{r, echo = FALSE, warning = FALSE, message = FALSE}

# Make a function to summarize the data
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}



```

# Plotting

### Ribbon plots

#### The number of RNA molecules by transcript type

```{r, echo = FALSE, warning = FALSE, message = FALSE}

#summarize the total number of RNAs
UTR_summ_total_RNA <- data_summary(quantTable, varname="Total_No_RNA", 
                    groupnames=c("Transcript", "X3UTR", "Cell_Stage"))

# plot number of RNA molecules by transcript type
ggplot(UTR_summ_total_RNA, aes(x=Cell_Stage, y=Total_No_RNA, group=Transcript, color=Transcript))+ 
  geom_ribbon(aes(ymin=Total_No_RNA-sd, ymax=Total_No_RNA+sd, fill = Transcript), alpha=0.1, linetype = 0) +
  facet_grid(. ~ X3UTR , scales = "free_y") +
  geom_line() + geom_point() +
  theme_linedraw() +
  theme(strip.text = element_text(face = "italic"))


```


#### The number of clusters by transcript type

```{r, echo = FALSE, warning = FALSE, message = FALSE}

# plot number of clusters by transcript type

UTR_summ_total_clusters <- data_summary(quantTable, varname="No_RNA_clusters", 
                    groupnames=c("Transcript", "X3UTR", "Cell_Stage"))


ggplot(UTR_summ_total_clusters, aes(x=Cell_Stage, y=No_RNA_clusters, group=Transcript, color=Transcript))+ 
  geom_ribbon(aes(ymin=No_RNA_clusters-sd, ymax=No_RNA_clusters+sd, fill = Transcript), alpha=0.1, linetype = 0) +
  facet_grid(. ~ X3UTR , scales = "free_y") +
  geom_line() + geom_point() +
  theme_linedraw() +
  theme(strip.text = element_text(face = "italic"))

```


#### The fraction of RNAs in clusters by transcript type

```{r, echo = FALSE, warning = FALSE, message = FALSE}

# plot fraction_RNA_in_clusters by transcript type


UTR_summ_fraction <- data_summary(quantTable, varname="fraction_RNA_in_clusters", 
                    groupnames=c("Transcript", "X3UTR", "Cell_Stage"))


ggplot(UTR_summ_fraction, aes(x=Cell_Stage, y=fraction_RNA_in_clusters, group=Transcript, color=Transcript))+ 
  geom_ribbon(aes(ymin=fraction_RNA_in_clusters-sd, ymax=fraction_RNA_in_clusters+sd, fill = Transcript), alpha=0.1, linetype = 0) +
  facet_grid(. ~ X3UTR , scales = "free_y") +
  geom_line() + geom_point() +
  theme_linedraw() +
  theme(strip.text = element_text(face = "italic"))


```


#### The RNA content of clusters by transcript type

```{r, echo = FALSE, warning = FALSE, message = FALSE}

# summarize RNA content of clusters by transcript type
UTR_summ_total_content <- data_summary(quantTable, varname="Average_No_RNA_per_Cluster", 
                    groupnames=c("Transcript", "X3UTR", "Cell_Stage"))

# summarize RNA content of clusters by transcript type
ggplot(UTR_summ_total_content, aes(x=Cell_Stage, y=Average_No_RNA_per_Cluster, group=Transcript, color=Transcript))+ 
  geom_ribbon(aes(ymin=Average_No_RNA_per_Cluster-sd, ymax=Average_No_RNA_per_Cluster+sd, fill = Transcript), alpha=0.1, linetype = 0) +
  facet_grid(. ~ X3UTR , scales = "free_y") +
  geom_line() + geom_point() +
  theme_linedraw() +
  theme(strip.text = element_text(face = "italic"))



```

### For figure plot

This plot will illustrate _cpg-2_ and _nos-2_ that are in the main plot (Fig2F, Fig2H)

```{r, echo = FALSE, warning = FALSE}

#Filter for cpg-2 and nos-2 that will be in the main figure
cpgnos2_summTotalSubset <- UTR_summ_total_content[which(UTR_summ_total_content$X3UTR == "cpg-2" | UTR_summ_total_content$X3UTR ==  "nos-2"),]

p_cpg2_nos2_AveMolPlot <- ggplot(cpgnos2_summTotalSubset, aes(x=Cell_Stage, y=Average_No_RNA_per_Cluster, group=Transcript, color=Transcript, shape = Transcript))+ 
  geom_ribbon(aes(ymin=Average_No_RNA_per_Cluster-sd, ymax=Average_No_RNA_per_Cluster+sd, fill = Transcript), alpha=0.1, linetype = 0) +
  facet_grid(. ~ X3UTR , scales = "free_y") +
  ylab("mRNA molecules per cluster") +
  geom_line() + geom_point() +
  theme_linedraw() +
  scale_fill_manual(values=c("magenta4", "green3", "magenta4")) +
  scale_color_manual(values=c("magenta4", "green3", "magenta4")) +
  scale_shape_manual(values=c(15, 17, 18, 6)) +
  theme(strip.text = element_text(face = "italic"), panel.grid.minor = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.text.y = element_text(angle=90), axis.text.x = element_text(angle=90)) 

p_cpg2_nos2_AveMolPlot

#Set date
date <- format(Sys.time(), "%y%m%d")
#set filename
filename <- paste("../3_output/", date, "_output_plots_Fig4FH.pdf", sep = "")


#export pdf
pdf(file = filename, width = 4, height = 2.5)
grid.arrange(p_cpg2_nos2_AveMolPlot)
dev.off()
dev.off()

#Report
print("The following output plot was saved: ", quote = FALSE)
print(filename, quote = FALSE)
```


### P value calculations

Perform Benjamini Hochberg adjusted t-tests on cpg-2 granule size v. neongreen::cpg-2 3'UTR granule size.

```{r, echo = FALSE, warning = FALSE}

##### Perform t-tests on cpg-2 v. neongreen::cpg-2 3'UTR localization to granules ########

# select cpg-2 data
cpg2QuantTable <- quantTable%>%
  filter(X3UTR == "cpg-2")
cpg2QuantTable$Transcript <- factor(cpg2QuantTable$Transcript, levels = c("cpg-2", "mNG"))

# summarize cpg-2 data
cpg2SummaryTable <- cpg2QuantTable%>%
  dplyr::group_by(Cell_Stage, Transcript) %>%
  dplyr::summarize(value = list(Average_No_RNA_per_Cluster)) %>%
  spread(Transcript, value)

# Perform t-tests
cpg2_pvalues <- data.frame()
for (i in 1:4){
  cpg2_pvalues[i,1] <- t.test(cpg2SummaryTable[[i,2]], cpg2SummaryTable[[i,3]])$p.value
  rownames(cpg2_pvalues)[i] <- as.character(cpg2SummaryTable$Cell_Stage[i])
}
colnames(cpg2_pvalues) <- c("p.value")

# Perform multiple test corrections using the Benjamini Hochberg method
cpg2_pvalues <- cbind(cpg2_pvalues, BH_padjust = p.adjust(cpg2_pvalues$p.value, method = "BH"))

# Report p values
cpg2_pvalues
```

Perform Benjamini Hochberg adjusted t-tests on nos-2 granule size v. neongreen::nos-2 3'UTR granule size.

```{r, echo = FALSE, warning = FALSE}

##### Perform t-tests on nos-2 v. neongreen::nos-2 3'UTR localization to granules ########

# select nos-2 data
nos2QuantTable <- quantTable%>%
  filter(X3UTR == "nos-2")
nos2QuantTable$Transcript <- factor(nos2QuantTable$Transcript, levels = c("nos-2", "mNG"))

# summarize nos-2 data
nos2SummaryTable <- nos2QuantTable%>%
  dplyr::group_by(Cell_Stage, Transcript) %>%
  dplyr::summarize(value = list(Average_No_RNA_per_Cluster)) %>%
  spread(Transcript, value)

# Perform t-tests
nos2_pvalues <- data.frame()
for (i in 1:4){
  nos2_pvalues[i,1] <- t.test(nos2SummaryTable[[i,2]], nos2SummaryTable[[i,3]])$p.value
  rownames(nos2_pvalues)[i] <- as.character(nos2SummaryTable$Cell_Stage[i])
}
colnames(nos2_pvalues) <- c("p.value")

# Perform multiple test corrections using the Benjamini Hochberg method
nos2_pvalues <- cbind(nos2_pvalues, BH_padjust = p.adjust(nos2_pvalues$p.value, method = "BH"))

# Report p values
nos2_pvalues

```

Next, I'll filter for chs-1 and clu-1. These will be in a supplemental figure: Figure4 Supplement 1.

```{r, echo = FALSE, warning = FALSE}

#Filter for chs-1 and clu-1 that will be in the main figure
chsclu_summTotalSubset <- UTR_summ_total_content[which(UTR_summ_total_content$X3UTR == "chs-1" | UTR_summ_total_content$X3UTR ==  "clu-1"),]

chsclu_summTotalSubset <- replace(chsclu_summTotalSubset, is.na(chsclu_summTotalSubset), 0)

p_chs1clu1_AveMolPlot <- ggplot(chsclu_summTotalSubset, aes(x=Cell_Stage, y=Average_No_RNA_per_Cluster, group=Transcript, color=Transcript, shape = Transcript))+ 
  geom_ribbon(aes(ymin=Average_No_RNA_per_Cluster-sd, ymax=Average_No_RNA_per_Cluster+sd, fill = Transcript), alpha=0.1, linetype = 0) +
  facet_grid(. ~ X3UTR , scales = "free_y") +
  ylab("mRNA molecules per cluster") +
  geom_line() + geom_point() +
  theme_linedraw() +
  scale_fill_manual(values=c("magenta4", "magenta4", "green3" )) +
  scale_color_manual(values=c("magenta4", "magenta4", "green3")) +
  scale_shape_manual(values=c(15, 17, 18, 6)) +
  theme(strip.text = element_text(face = "italic"), panel.grid.minor = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.text.y = element_text(angle=90), axis.text.x = element_text(angle=90)) 

p_chs1clu1_AveMolPlot

#Set date
date <- format(Sys.time(), "%y%m%d")
#set filename
filename <- paste("../3_output/", date, "_output_plots_Fig4Supp1.pdf", sep = "")


#export pdf
pdf(file = filename, width = 4, height = 2.5)
p_chs1clu1_AveMolPlot
dev.off()
dev.off()

#Report
print("The following output plot was saved: ", quote = FALSE)
print(filename, quote = FALSE)


```

### P value calculations

Perform Benjamini Hochberg adjusted t-tests on chs-1 v. neongreen::chs-1 3'UTR granule size.

```{r, echo = FALSE, warning = FALSE}

##### Perform t-tests on chs-1 v. neongreen::chs-1 3'UTR localization to granules ########

# select chs-1 data
chs1QuantTable <- quantTable%>%
  filter(X3UTR == "chs-1")
chs1QuantTable$Transcript <- factor(chs1QuantTable$Transcript, levels = c("chs-1", "mNG"))

# summarize chs-1 data
chs1SummaryTable <- chs1QuantTable%>%
  dplyr::group_by(Cell_Stage, Transcript) %>%
  dplyr::summarize(value = list(Average_No_RNA_per_Cluster)) %>%
  spread(Transcript, value)

# Perform t-tests
chs1_pvalues <- data.frame()

for (i in 1:2){
  chs1_pvalues[i,1] <- t.test(chs1SummaryTable[[i,2]], chs1SummaryTable[[i,3]])$p.value
  rownames(chs1_pvalues)[i] <- as.character(chs1SummaryTable$Cell_Stage[i])
}
colnames(chs1_pvalues) <- c("p.value")


# Perform multiple test corrections using the Benjamini Hochberg method
chs1_pvalues <- cbind(chs1_pvalues, BH_padjust = p.adjust(chs1_pvalues$p.value, method = "BH"))

# Report p values
chs1_pvalues
```

Perform Benjamini Hochberg adjusted t-tests on clu-1 v. neongreen::clu-1 3'UTR granule size.

```{r, echo = FALSE, warning = FALSE}

##### Perform t-tests on clu-1 v. neongreen::clu 3'UTR localization to granules ########

# select clu-1 data
cluQuantTable <- quantTable%>%
  filter(X3UTR == "clu-1")
cluQuantTable$Transcript <- factor(cluQuantTable$Transcript, levels = c("clu-1", "mNG"))

# summarize chs-1 data
clu1SummaryTable <- cluQuantTable%>%
  dplyr::group_by(Cell_Stage, Transcript) %>%
  dplyr::summarize(value = list(Average_No_RNA_per_Cluster)) %>%
  spread(Transcript, value)

# Perform t-tests
clu1_pvalues <- data.frame()

for (i in 1:4){
  clu1_pvalues[i,1] <- t.test(clu1SummaryTable[[i,2]], clu1SummaryTable[[i,3]])$p.value
  rownames(clu1_pvalues)[i] <- as.character(clu1SummaryTable$Cell_Stage[i])
}
colnames(clu1_pvalues) <- c("p.value")

# Perform multiple test corrections using the Benjamini Hochberg method
clu1_pvalues <- cbind(clu1_pvalues, BH_padjust = p.adjust(clu1_pvalues$p.value, method = "BH"))

# Report p values
clu1_pvalues
```


# Get versions
```{r, echo = TRUE}
sessionInfo()
```
